---
title: "Problem Set 1"
subtitle: "Intro to `R`"
author: "[YOUR NAME]"
institute: "Villanova University"
date: "Due Date: 2024-10-25"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps1.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps1.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using the `sc_debt.Rds` file from the course [github page](https://github.com/rweldzius/PSC7000_F2024/raw/main/Data/sc_debt.Rds).

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. 

This problem set is worth 5 total points. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compile the completed problem set and upload the PDF file to Blackboard on Friday by midnight. If you use OpenAI for help, choose to save your output as a PDF and then add this to the PDF for the problem set. Also note that I will not respond to Campuswire messages after 5PM on Friday, so don't wait until the last minute to get started!

**Good luck!**


## Question 0 [0 points]
*Require `tidyverse` and load the `sc_debt.Rds` data by assigning it to an object named `df`.*
```{r}
require() # Load tidyverse
df <- read_rds() # Load the dataset directly from github
```


## Question 1 [0.5 points]
*Which school has the lowest admission rate (`adm_rate`) and which state is it in (`stabbr`)?*
```{r}
df %>% 
  arrange() %>% # Arrange by the admission rate
  select() # Select the school name, the admission rate, and the state
```

> Write answer here

*Which are the top 10 schools by average SAT score (`sat_avg`)?*
```{r}
df %>%
  arrange() %>% # arrange by SAT scores in descending order
  select() %>% # Select the school name and SAT score
  print() # Print the first 12 rows (hint: there is a tie)
```

> Write answer here

## Question 3 [0.5 points]
*Create a new variable called `adm_rate_pct` which is the admissions rate multiplied by 100 to convert from a 0-to-1 decimal to a 0-to-100 percentage point.*

```{r}
df <- df %>% # Use the object assignment operator to overwrite the df object
  mutate() # Create the new variable adm_rate_pct
```

*Calculate the average SAT score and median earnings of recent graduates by state.*
```{r}
df %>%
  group_by() %>% # Calculate state-by-state with group_by()
  summarise(sat_avg = , # Summarise the average SAT
            earn_avg = ) # Summarise the average earnings
```

## Question 4 [0.5 points]
*Plot the average SAT score (x-axis) against the median earnings of recent graduates (y-axis) by school, and add the line of best fit. What relationship do you observe? Why do you think this relationship exists?*
```{r}
# INSERT CODE HERE
```

> Write answer here

## Question 5 [0.5 points]
*Research Question: Do students who graduate from smaller schools (i.e., schools with smaller student bodies) make more money in their future careers? Before looking at the data, write out what you think the answer is, and explain why you think so.*

> Write a few sentences here.

## Question 6 [0.5 points]
*Based on this research question, what is the outcome / dependent / $Y$ variable and what is the explanatory / independent / $X$ variable? Create the scatterplot of the data based on this answer, along with a line of best fit. Is your answer to the research question supported?*
```{r}
df %>%
  ggplot(aes(x = , # Put the explanatory variable on the x-axis
             y = )) +  # Put the outcome variable on the y-axis
  geom_point() + # Create a scatterplot
  geom_smooth() + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       x = '',
       y = '')
```

> Write a few sentences here.

## Question 7 [0.5 points]
*Does this relationship change by whether the school is a research university? Using the filter() function, create two versions of the plot, one for research universities and the other for non-research universities.*

```{r}
df %>%
  filter() %>% # Filter to non-research universities
  ggplot(aes(x = , # Put the explanatory variable on the x-axis
             y = )) +  # Put the outcome variable on the y-axis
  geom_point() + # Create a scatterplot
  geom_smooth() + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       subtitle = '', 
       x = '',
       y = '')

df %>%
  filter() %>% # Filter to research universities
  ggplot(aes(x = , # Put the explanatory variable on the x-axis
             y = )) +  # Put the outcome variable on the y-axis
  geom_point() + # Create a scatterplot
  geom_smooth() + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       subtitle = '', 
       x = '',
       y = '')
```


## Question 8 [0.5 points]
*Instead of creating two separate plots, color the points by whether the school is a research university. To do this, you first need to modify the research_u variable to be categorical (it is currently stored as numeric). To do this, use the mutate command with `ifelse()` to create a new variable called `research_u_cat` which is either "Research" if `research_u` is equal to 1, and "Non-Research" otherwise.*
```{r}
df <- df %>%
  mutate(research_u_cat = ifelse()) # Create a labeled version of the research_u variable

df %>%
  ggplot(aes(x = , # Put the explanatory variable on the x-axis
             y = , # Put the outcome variable on the y-axis
             color = )) + # Color the points by the new variable you created above
  geom_point() + # Create a scatterplot
  geom_smooth() + # Add line of best fit
  labs(title = '', # give the plot meaningful labels to help the viewer understand it
       x = '',
       color = '',
       y = '')
```

## Question 9 [1 point]
*Write a short paragraph discussing your findings. What do you think is going on in these data?*

> Write a few sentences here


