---
title: "Problem Set 2"
subtitle: "Data Wrangling & Univariate Analysis"
author: "[YOUR NAME]"
institute: "Villanova University"
date: "Due Date: 2024-11-01"
output:
  html_document: default
---

```{r,include=F}
knitr::opts_chunk$set(error=TRUE)
```


## Getting Set Up

Open `RStudio` and create a new RMarkDown file (`.Rmd`) by going to `File -> New File -> R Markdown...`.
Accept defaults and save this file as `[LAST NAME]_ps2.Rmd` to your `code` folder.

Copy and paste the contents of this `.Rmd` file into your `[LAST NAME]_ps2.Rmd` file. Then change the `author: [Your Name]` to your name.

We will be using the `MI2020_ExitPoll.Rds` file from the course [github page](https://github.com/rweldzius/PSC7000_F2024/raw/main/Data/MI2020_ExitPoll.Rds) for questions 1-4 and the `nba_players_2018.Rds` file from the course [github page](https://github.com/rweldzius/PSC4175_F2024/raw/main/Data/nba_players_2018.Rds) for questions 5-8.

All of the following questions should be answered in this `.Rmd` file. There are code chunks with incomplete code that need to be filled in. To submit, compile (i.e., `knit`) the completed problem set and upload the PDF file to Dropbox on Friday by midnight. Instructions for how to compile the output as a PDF can be found in [Problem Set 0](https://github.com/rweldzius/PSC7000_F2024/raw/main/Psets/psc7000_pset_0.pdf). 

This problem set is worth 5 total points. The point values for each question are indicated in brackets below. To receive full credit, you must have the correct code. In addition, some questions ask you to provide a written response in addition to the code.

You will be deducted 1 point for each day late the problem set is submitted, and 1 point for failing to submit in the correct format (PDF).

You are free to rely on whatever resources you need to complete this problem set, including lecture notes, lecture presentations, Google, your classmates...you name it. However, the final submission must be complete by you. There are no group assignments. To submit, compile the completed problem set and upload the PDF file to Drobox on Friday by midnight. If you use OpenAI for help, choose to save your output as a PDF and then add this to the PDF for the problem set. Also note that I will not respond to Campuswire messages after 5PM on Friday, so don't wait until the last minute to get started!

**Good luck!**


## Question 0.A
Require `tidyverse` and an additional package called `labelled` (remember to `install.packages("labelled")` if you don't have it yet) and load the `MI2020_ExitPoll.Rds` data to an object called `MI_raw`. (Tip: use the `read_rds()` function with the link to the raw data.)

```{r}
require()
require()
MI_raw <- read_rds('') 
```


## Question 1 [0.5 point]
*What is the unit of analysis in this dataset? How many variables does it have? How many observations?*

> Write answer here

## Question 2 [0.5 point]
*This has too much information that we don't care about. Create a new object called `MI_clean` that contains only the following variables:*

* AGE10
* SEX
* PARTYID
* EDUC18
* PRSMI20
* QLT20
* LGBT
* BRNAGAIN
* LATINOS
* QRACEAI
* WEIGHT

*and then list which of these variables contain missing data recorded as `NA`. How many respondents were not asked certain questions?*
```{r}
MI_clean <- MI_raw %>% 
  select(...) # Select the requested variables

summary(...) # Identify which have missing data recorded as NA
```

> Write answer here

## Question 3 [0.5 point]
*Are there* **unit non-response** *data in the `PRSMI20` variable? If so, how are they recorded? What about the `PARTYID` variable? How many people refused to answer both of these questions?*

```{r}
MI_clean %>%
  count(...) # Tip: use count() function to look at your variables.
```

> Write answer here.

## Question 4 [0.5 points]
*Let's create a new variable called `preschoice` that converts `PRSMI20` to a character. To do this, install the `labelled` package if you haven't already, then use the `to_character()` function from the `labelled` package. Now `count()` the number of respondents who reported voting for each candidate. How many respondents voted for candidate Trump in 2020? How many respondents refused to tell us who they voted for?*
```{r}
MI_clean <- MI_clean %>%
  mutate(preschoice = ...) # Convert to character

MI_clean %>%
  count(...)
```

> Write answer here


## Question 5 [0.5 point]
What proportion of women supported Trump?

```{r}
# Women Trump supporters
MI_clean %>%
  drop_na(...) %>% # Drop any missing values for preschoice
  filter(...) %>% # Filter to only women
  count(...) %>% # Count the number of women who supported each candidate
  mutate(share = ...) # Calculate the proportion of women who supported Trump

# Alternative approach
MI_clean %>%
  drop_na(...) %>% # Drop any missing values for preschoice
  mutate(trumpSupp = ifelse(...)) %>% # Create "dummy" variable for whether the person voted for Trump or not that is either 1 (they voted for Trump) or 0
  group_by(...) %>% # Group by gender
  summarise(share = mean(trumpSupp)) # Calculate proportion who supported Trump
```

> Write answer here.


## Extra Credit 1 [0.5 point]
*Among women, which age group sees the highest support for Trump? To answer, you will need to calculate the proportion of women who supported Trump by age-group to determine which age-group had the highest Trump support among women. You will need to clean the AGE10 variable before completing this problem, just like we did with the PRSMI20 variable. Call the new variable "Age". HINT: to make your life easier (and not write a 10-level nested ifelse() function), try asking ChatGPT for help with this prompt: "I have a labelled variable in R that I want to convert to text. How can I do this?"*

```{r}
# Insert code here.
```

> Write answer here

## Question 0.B
*Require `tidyverse` and an additional package called `haven` (remember to `install.packages("haven")` if you don't have it yet; once you've installed it, remember to either delete the command or comment it out with #), and then load the [`nba_players_2018.Rds`](https://github.com/rweldzius/PSC7000_F2024/raw/main/Data/nba_players_2018.Rds?raw=true') data to an object called `nba`.*

```{r}
# Insert code here.
```


## Question 6 [0.5 point]
*Plot the distribution of field goals attempted by all NBA players in the 2018-2019 season. Explain why you chose the visualization that you did. Then add a vertical line indicating the mean and median number of points in the data. Color the median line blue and the mean line red. Why is the median lower than the mean?*
```{r}
nba %>%
  ggplot(...) + # Put the fga variable on the x-axis of a ggplot.
   geom_...() + # Choose the appropriate geom function to visualize.
  labs(...) + # Add labels
      geom_vline(...) + # Median vertical line (blue)
      geom_vline(...) # Mean vertical line (red)
```

> Write answer here.

## Question 7 [0.5 point]
*Now examine the `country` variable. Visualize this variable using the appropriate `geom_...`, and justify your reason for choosing it. Tweak the plot to put the country labels on the y-axis, ordered by frequency. Which country are most NBA players from? What is weird about your answer, and what might explain it? *

```{r}
nba %>%
  count(...) %>% # count the number of players by country
  ggplot(...) + # place the country on the y-axis, reordered by the number of players. Put the number of players on the x-axis
  geom_...(...) + # Choose the best geom
  labs(...) # Add labels
```

> Write answer here

## Question 8 [0.5 point]
*Let's pretend we are consulting for an NBA organization. The owner and GM tell us they are interested in the relationship between the player's age (`agePlayer`) and the amount of points they score (`pts`). Please answer the following research question and provide a theory supporting your answer: "Do older NBA players score more points than younger players?"*

> Write answer here

## Question 9 [0.5 point]
*Based on your answer above, what is the outcome / dependent / $Y$ variable and what is the explanatory / independent / $X$ variable? Why?*

> Write answer here

*Create a univariate visualization of both the $X$ and $Y$ variables. Choose the best `geom_...()` based on the variable type, and make sure to label your plots!*

```{r}
# X variable
nba %>%
  ggplot(...) + # Put the X variable on the x-axis
  geom_...() +  # Choose the best geom given the variable type (make sure to look at it if you aren't sure)
  labs(...)     # Add labels

# Y variable
nba %>%
  ggplot(...) + # Put the Y variable on the x-axis
  geom_...() +  # Choose the best geom given the variable type (make sure to look at it if you aren't sure)
  labs(...)     # Add labels
```

## Question 10 [0.5 point]
*Now analyze the data by creating a multivariate visualization that shows the relationship between age and points. Add a STRAIGHT line of best fit with `geom_smooth()`.*

```{r}
nba %>%
  ggplot(...) + # Put the X variable on the x-axis, and the Y variable on the y-axis
  geom_...() +  # Choose the best geom given both variable types
  geom_smooth(...) + # Add a STRAIGHT line of best fit
  labs(...)     # Add labels
```

*Based on your analysis, does the data support or reject your hypothesis from Question 3?*

> Write answer here

## Extra Credit 2 [0.5 point]
*Let's look for evidence of a "curvelinear" relationship between player age and points scored. To do so, first calculate the average points scored by age. Then plot this relationship using a multivariate visualization. Add a line of best fit with `geom_smooth()` but DON'T use `method = "lm"`. What do you conclude? Why?*

```{r}
# INSERT CODE HERE
```

> Write answer here
